MS Depression Paper, post replication, started 2022 09 21

### 2023-05-17 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
added corpus callosum sensitivity analyses

@@

```{r sensitivity analyses}

@@

>>>>

#scatter for Taki
scatter_net <- ggplot(newdf,  aes(x=dep_net_group, y=net_values)) +  geom_jitter(shape=16, position=position_jitter(0.2))

scatter_dx <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum)) +  geom_jitter(shape=16, position=position_jitter(0.2))

@@

#### tercile/quintile cutoffs
```{r sensitivity_analyses_diff_depression_cutoffs}

@@

#### dropping corpus callosum
```{r sensitivity_analysis_no_cc}
## compare disease within and outside depression network, between diagnoses, and interactions

## volume of corpus collosum
volume_of_healthy_corpus_callosum = fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[fascicle_volumes_dep_network_no_cranial_nerves$fascicle == "CC"]

#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
dep_and_healthy_groups_for_ICD_analysis_no_cc <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>%

  #sum of total volume lost and proportion of volume lost 
  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix[fascicle_names_w_vol_suffix != "CC_vol"])))) %>%
  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/(total_volume_all_full_fascicles-volume_of_healthy_corpus_callosum)) %>%

  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves[dep_network_names_top_quartile_no_cranial_nerves != "CC_vol"])))) %>%
  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /(dep_network_total_volume_top_quartile_no_cranial_nerves-volume_of_healthy_corpus_callosum)) %>%

  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles/bottom_75_total_volume_no_cranial_nerves) %>%

#clean up data frame
df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis_no_cc %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#diagnosis: Dep > nondep, P= 0.013, T 2.49, Estimate 0.06 ; Network dep>nondep P = 9.48*10^-5, T = 3.95, estimate 0.03; intx  P=.03, T = -2.12, Estimate = -0.02
diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
summary(diagnosis_by_network_interaction_lmer)

#Diagnosis: MS+ > MS- 0.01-0.10; Network dep>nondep 0.02-0.05; intx: -0.04--0.002 
confint(diagnosis_by_network_interaction_lmer)

#### Volume x Effect size analysis

fascicle_dep_vs_healthy_effsize <- sapply(fascicle_names_w_vol_suffix[fascicle_names_w_vol_suffix != "CC_vol"], function(fascicle) 
{

  text_for_eval <- paste0("rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis_no_cc, formula = ", fascicle, " ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = \"perc\", nboot = 1000)$effsize")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_effsize) <- fascicle_names_w_vol_suffix[fascicle_names_w_vol_suffix != "CC_vol"]

df_fascicle_dep_mask_volume_overlap_effsize_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle[fascicle_volumes_dep_network_no_cranial_nerves$fascicle != "CC"], fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map[fascicle_volumes_dep_network_no_cranial_nerves$fascicle != "CC"], df_effsize_values)
names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")

# plot fascicle overlap x Eff size value
#statistic, p=0.0002, T =3.865 , eff size (adjusted R^2 = 0.16)
dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )
summary(dep_overlap_x_dep_vs_healthy_vol_effsize)

scatter_effsize_val_by_dep_net_overlap_log <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize, color=dep_vs_healthy_effsize)) +
  geom_point(size=2) + 
  scale_color_gradient(low = "#FFFFFF", high = "#ff00ff") +
  geom_smooth(formula = y ~ x, method=lm, color="black")+
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(0, 10, 100, 1000, 10000, 100000)) + 
  xlab("Volume of Fascicle in Depression Network, excluding CC") + 
  ylab("Effect Size (r)") + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=16, color = "black"),
        axis.title = element_text(size=18, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank(),
        legend.position = "none",
        aspect.ratio = 4/4) + 
  annotate("text", x=1.6, y=0.14, size = 5, 
           label = paste0("P = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize)$coef[2,4],4)))

#Final data: 
#Main effect of network Dep > Nondep (Estimate = 0.03, 95% CI 0.02-0.05 , P= 9.48*10^-5)
#Main effecet of diagnosis MS+Depression > MS-Depression (Estimate = 0.06, 95% CI 0.01-0.10 , P = 0.013)
#Main effect of interaction (Estimate = 0.02, 95% CI 0.002 -0.04 , P=0.03)
